### 2022-08-30 paper/icse23/approach.tex

\begin{figure*}[t]
  \centering
  \includegraphics[width=\linewidth]{figs/overview.pdf}
  \caption{\OverviewFigCaption}
\end{figure*}

@@

  We design and implement a new NLP model testing method,
  \emph{Specification- and Syntax-based Linguistic Capability Testing
    (\tool{})}, that automatically generate test cases with oracles to
  test the robustness of sentiment analysis models. \tool{} addresses
  all three challenges discussed above.

@@

% We present \tool{}, a new \lc test case generation
% tool, that satisfies all of these criteria.  \sw{I could not summarize
%   a cohesive idea that drives our design. Leaving it here to fill in.
%   Also, I felt my writing below still lacks justification for some
%   design choices (e.g., why we do the differentiation).}
Figure~\ref{fig:overview} shows the overview of \tool{}, which
consists of two phases.  The \emph{specification-based seed
  generation} phase performs rule-based searches from a real-world
dataset and template-based transformation to obtain the initial seed
\sents.  The search rules (e.g., search for \neu \sents that do not
include any \pstv or \ngtv words) and transformation templates (e.g.,
negating a sentence) are defined in the \emph{\lc specifications},
which guarantee that each resulting seed conforms to a specific \lc
({\bf C2}) and is labelled correctly ({\bf C3}).

The \emph{syntax-based \sent expansion} phase expands the seed \sents
with additional syntactic elements (i.e., words) to cover many
real-world syntactic structures ({\bf C1}). It first performs a syntax
analysis to identify the \pos (PoS) tags that can be inserted to each
seed, by comparing the PoS parse trees between the seed \sent and many
other \sents from a large reference dataset. Each identified tag is
inserted into the seed as a \emph{mask}. It then uses an NLP
recommendation model (i.e., BERT \cite{}) to suggest possible
words. If a resulting \sent is validated to be consistent with the
specification which additionally defines the rules for expansion
(e.g., the expanded word should be \neu), {\bf C2} and {\bf C3} are
still satisfied. Last, because some validated \sents may include
unacceptable suggested words given the context, we use a heuristic
(i.e., the confidence score from the NLP recommendation model) to
select the more realistic context-aware expanded \sents into \tool{}'s
test suite.

@@

\subsection{Specification-based Seed Generation}
The seed generation phase of \tool starts by searching \sents in a
real-world dataset that match the rules defined in the \lc
specification, and then transforming the matched \sents using
templates to generate seed \sents that conform to individual \lcs. The
reasons for this design choice are twofold.  First, while generally
judging which \lc any \sent falls into and which label it should have
is infeasible, there exist simple rules and templates to allow
classifying the resulting \sents into individual \lcs and with the
correct labels, with high confidence.  This enables us to test each
\lc individually.  Second, searching from a real-world dataset ensures
that the \sents used as test cases for testing \lcs are realistic and
diverse. The diverse test cases are more likely to achieve a high
coverage of the target model's functionality in each \lc, thus
detecting more errors.
%
In this phase, \tool first search and selects \sents applicable to the
\lc in a given real-world dataset with search rules. In case that the
search rules only fulfill portion of the \lc specifications, the
selected \sents are not yet appropriate to become seed, we transform
the selected \sents into seed \sents using the heuristic templates.
%
Table~\ref{tab:specification} shows the search rules and the
transformation templates of all 11 linguistic capabilities we
implemented in \tool. The first column shows the \lc type and its
description, and the second column shows the search rule and
transformation template used in each \lc. For LC1 and LC2, the NLP
models are evaluated in the scope of short \sents with selective
sentiment words. It does not require any transformation because the
search rule alone is sufficient to conform to the \lcs. On the other
hand, search rules of LC3 to LC11 are not enough to match their \lc
specification, thus \tool uses heuristic templates to conform the
searched \sents to the \lc. For example, in LC3's transformation
template, the searched \sents become seeds by preturbing them with the
defined templates.  In LC4's transformation template, the searched
demonstrative sentences are negated.
% \sw{@Jaeseong: revise the rest of 3.1 based on
%   Table~\ref{tab:specification}. I commented out the old text but it
%   is still in the tex file.} \jl{I revised and added the explanation}

@@

Algorithm \ref{alg:diff} shows how masks are identified for each seed
\sent.  It takes the parse trees of the seeds, generated by the
Berkeley Neural
Parser~\cite{kitaev2018seedparser,kitaev2019seedparser}, and a
reference context-free grammar (CFG) from the Penn Treebank corpus
dataset \cite{mitchell1993treebank} as inputs.  The reference CFG is
learned from a large dataset \cite{nltkTreebankCorporaWebPage} that is
representative of the distribution of real-world language usage.  The
algorithm identifies the discrepancy between the seed syntax and the
reference grammar to decide how a seed can be expanded.
%This phase builds \pcfg from the reference corpus. \cfg
%is constructed by parsing \sents in the corpus and extracting
%\prodrs. In addition, the probability of \cfg is estimated by
%its frequency over corpus. The output of this phase is the constructed
%\pcfg, and it is compared with seed parse trees. For our
%implementation, we build the \pcfg from the Penn Treebank corpus dataset.

@@

For each production of in each seed's parse tree (lines 3 and 4), we
extract its non-terminal at the left-hand-side (line 5), $s\_lhs$, and
the grammar symbols at the right-hand-side (line 6), $s\_rhs$. In line
7, the algorithm iterates through all productions in the reference
context-free grammar and match these that have the same non-terminal
at the left-hand-side as $s\_lhs$.  The right-hand-side of each
matched production is called $r\_rhs$.  If $s\_rhs$ consists of a
subset of the grammar symbols in $r\_rhs$ (line 8), the additional
symbols in the $r\_rhs$ are inserted as masks in the parse tree of
seed \sent, in their respective positions in the expanded production.
The left to right traversal of the leaves of an expanded parse tree
forms a masked \sent.  Lastly, due to the inefficient cost of
accessing full list of the masked \sents, we randomly select $k$
masked \sents for the next \sent expansion and validation phase when
the masked \sents are more than maximum number of masked \sents. The
random sampling is unbiased approach since it gives same chance to be
chosen. Thus, the random sample becomes representative of the
population of the masked \sents, and it efficiently shows the
usefulness of the \tool.

\begin{figure*}[t]
  \centering
  \includegraphics[scale=0.7]{figs/running_example.pdf}
  \caption{\RunningExCaption}
%  Expansion of the seed \sent ``Or
%both.''. For a \prodr in seed ``NP->[DT]'' on the left, the \prodr of
%``NP->[DT, NNS]'' is found in reference. Thus, the component NNS can be expanded in the
%seed and ``NP->[DT]'' is replaced with ``NP->[DT, NNS]'' and it generates new
%expanded \sent on the right.
\end{figure*}

\paragraph{Running example.} Figure~\ref{fig:ExpEx} shows an example using Algorithm \ref{alg:diff}
to generate a masked \sent. The \sent ``Or both." is a seed of \lc of
\SareqExOne.  The tree on the left shows the parse tree of this seed;
it consists of two productions: ``FRAG->[CC, NP, .]" and ``NP->[DT]".
When matching the left-hand-side non-terminal of the second production
(i.e., ``NP") in the reference CFG, we found that it includes a
production ``NP->[DT, NNS]" which has an additional symbol ``NNS" on
the right-hand-side.  The algorithm thus expands the parse tree with
this symbol, shown in the second tree.  The masked \sent ``Or both
\{MASK\}." is the result of the left-to-right traversal of this
expanded parse tree.

%Given the seed parse tree and reference \pcfg, production
%differentiation phase suggests structural expansion candidates on the
%seed input. this phase aims to analyze which structural components
%and where they can be added into the seed structure for its
%expansion. To do so, we explore reference \prodrs comparing it with
%each \prodr used in seed input. This results in the phase described
%in Algorithm~\ref{code:ProdDiffAlg}. For each production rule in seed
%inputs ($seed\_prod$), it searches production rules in reference
%($ref\_prod$) which it has same non-terminal on the \lhs
%($seed\_prod.lhs==ref\_prod.lhs$) and superset of \rhs of the seed
%production rule ($seed\_prod.rhs \subset ref\_prod.rhs$).  As we
%assume that the reference \cfg is built from real world data
%distribution, the elements in the complement set ($ref\_prod.rhs -
%seed\_prod.rhs$) become an expansion candidate which can be expanded
%from the $seed\_prod.rhs$ found in real world. In addition, the
%measure of how consistent the production rule is with the given seed
%structure is given in its probability of the reference production
%rule ($ref\_prob$) multiplied by that of parents of $seed\_prod$. The
%expansion candidate consists of terminal or nonterminal symbols. When
%there is a phrase-level or clause-level nonterminal symbol, \eg noun
%phrase, it needs to be expanded and replaced with word-level
%nonterminal or terminal symbols to generate the expansion
%candidate. The number of feasible replacement is unbounded because of
%its high degree of freedom. Therefore, in this work, we focus on the
%expansion candidate with only the word-level nonterminal or terminal
%symbols for the effectiveness of \Model. Lastly, the expanded
%component is replaced with the mask token for the next phase. The
%example of the expansion is illustrated in
%Figure~\ref{fig:ExpEx}. The ``NP->[DT]'' is queried into reference,
%and ``NP->[DT,NNS]'' is identified as its expansion candidate since
%the \rhs of ``NP->[DT,NNS]'' is superset of that of ``NP->[DT]''. the
%component of NNS is replaced with mask token in
%\sent-level. Therefore, the ``Or both \{MASK\}.'' is suggested for
%the next phase.

@@

In this phase, the words to fill in the masks in the masked \sents are
suggested by the BERT pretrained model~\cite{devlin2019bert}.  The
BERT is a transformer-based \nl model. It is pretrained on two tasks
of masked token prediction and next sentence prediction. As a result
of the training process, the BERT model suggests word for the mask
token according to its surrounding context in \sent. For each masked
token, multiple words are suggested ranked by their confidence scores.
% \sw{Algorithm 1 does not require we only have one mask in the masked
%   \sent. Can the model suggest multiple words at the same time?}
% \jl{yes. it can predict multiple masked tokens at the same time.}
% \jl{I added the explanation of BERT}
%
Because BERT model is not aware of the \lc specification and the
grammar symbol in the expanded parse tree, an expanded \sent using the
suggested words may no longer satisfy the \lc
specification. Therefore, we perform validation on the suggested words
and only accept them if the following three criteria are met.

First, the PoS tag of the suggested word must match the PoS tag of the
expanded symbol in the parse tree. For the example in
Figure~\ref{fig:ExpEx}, the masked symbol is a ``NNS" (i.e., plural
noun); thus, the suggested word must also be a ``NNS". In this work,
we use \spacy, a free open-source library for \nlp, for extracting PoS
tags for each suggested word.

Second, it is required that the sentiment of the expanded \sent
becomes the same as the seed \sent. To ensure this, the suggested
words must be \neu.

Third, we additionally verify that the expanded \sents satisfied the
same search rules for the seed \sent in LC1 and LC2. This criteria
cannot be applied to other linguistic capabilities because they have
additional transformation templates.

In this work, out of the BERT suggested words, we randomly selects
words that matches their PoS tag of the expanded symbol in the parse
tree. The words that meets the second and third criteria are finally
employed for the expanded \sents.
% Our goal for generating the expanded \sents is to use them for
% evaluating the \sa models on the associated \lc in addition to the
% seed \sent. It is only achieved when the expanded \sents are also
% met with the same search rules for the \lc. For LC1 as an example,
% the expanded \sent must still conform to the specification of the
% seed's \lc specification. Therefore, the expanded \sents are
% required to be short and to only have \neu adjectives and \nns.
% \sw{Say why we use the third criteria only for LC1 and LC2.}  \jl{I
% added the explanation}

>>>>

\begin{figure*}[t]
  \centering
  \includegraphics[width=\linewidth]{figs/overview.pdf}
  \vspace{-6mm}
  \caption{\OverviewFigCaption}
\end{figure*}

@@

  We design and implement a new NLP model testing method,
\emph{Specification- and Syntax-based Linguistic Capability Testing
  (\tool{})}, that automatically generates test cases with oracles to assess the quality of sentiment analysis models. \tool{} addresses all three challenges discussed in Section~\ref{sec:intro}.

@@

% We present \tool{}, a new \lc test case generation
% tool, that satisfies all of these criteria.  \sw{I could not summarize
%   a cohesive idea that drives our design. Leaving it here to fill in.
%   Also, I felt my writing below still lacks justification for some
%   design choices (e.g., why we do the differentiation).}
Figure~\ref{fig:overview} shows the overview of \tool{}, which
consists of two phases.  The \emph{specification-based seed
  generation} phase performs rule-based searches from a real-world
dataset and template-based transformation to obtain the
initial seed \sents.  The search rules (e.g., search for \neu
\sents that do not include any \pstv or \ngtv words) and
transformation templates (e.g., negating a sentence) are defined in the \emph{\lc
  specifications} (Table~\ref{tab:specification}), which result in seed sentences that are most likely to conform
to a specific \lc ({\bf C2}) and is labelled
correctly ({\bf C3}).

The \emph{syntax-based \sent expansion} phase expands the seed
\sents with additional syntactic elements (i.e., words) to cover many real-world
syntactic structures ({\bf C1}). It first performs a syntax analysis
to identify the \pos (PoS) tags that can be inserted to each
seed, by comparing the PoS parse trees between the seed \sent and
many other \sents from a large reference dataset. Each identified
tag is inserted into the seed as a \emph{mask}. It then uses an NLP
recommendation model (i.e., BERT \cite{devlin2019bert}) to suggest possible
words. If a resulting \sent is validated to be consistent with the
specification which additionally defines the rules for expansion
(e.g., the expanded word should be \neu), {\bf C2} and {\bf C3} are
still satisfied.
% Last, because some validated \sents may include
% unacceptable suggested words given the context, we use a heuristic (i.e., the
% confidence score from the NLP recommendation model) to select the more
% realistic context-aware expanded \sents into \tool{}'s test suite.

@@

\subsection{Specification-based Seed Generation}
The seed generation phase of \tool starts by searching \sents in a
real-world dataset that match the rules defined in the \lc
specification, and then transforming the matched \sents using
templates to generate seed \sents that conform to individual
\lcs. The reasons for this design choice are twofold.  First, while
generally judging which \lc any \sent falls into and which label it
should have is infeasible, there exist simple rules and templates to
allow classifying the resulting \sents into individual \lcs and
with the correct labels, with high confidence.  This enables us to
test each \lc individually.  Second, searching from a real-world
dataset ensures that the \sents used as test cases for testing \lcs
are realistic and diverse. The diverse test cases are more likely to
achieve a high coverage of the target model's functionality in each
\lc, thus detecting more errors.
%
In this phase, \tool first searches and selects \sents applicable to
the \lc in a given real-world dataset with search rules. In case that
the search rules only fulfill portion of the \lc specifications, the
selected \sents are not yet ready to be used as seeds, we
transform the selected \sents into seed \sents using the
heuristic templates.
%
Table~\ref{tab:specification} shows the search rules and the
transformation templates of all 11 linguistic capabilities we
implemented in \tool. The first column shows the \lc type and its
description, and the second column shows the search rule and
transformation template used in each \lc. For LC1 and LC2, the NLP
models are evaluated in the scope of short \sents with selective
sentiment words. It does not require any
transformation because the search rules alone are sufficient to find sentences that conform
to the \lcs. On the other hand, search rules of LC3 to LC11 are not enough to match their \lc specification; thus, \tool additionally uses 
templates to transform the searched \sents to match the corresponding \lc. For example, in LC3's transformation template, the searched \sents are appended as part of a sentence to generate a seed.
In LC4's transformation template, the searched demonstrative sentences are negated.
% \sw{@Jaeseong: revise the rest of 3.1 based on
%   Table~\ref{tab:specification}. I commented out the old text but it
%   is still in the tex file.} \jl{I revised and added the explanation}

@@

Algorithm \ref{alg:diff} shows how masks are identified for each seed
\sent.  It takes the parse trees of the seeds, generated by the
Berkeley Neural
Parser~\cite{kitaev2018seedparser,kitaev2019seedparser}, and a
reference context-free grammar (CFG) from the Penn Treebank corpus
dataset~\cite{mitchell1993treebank} as inputs.  
This reference CFG was learned from a large dataset~\cite{nltkTreebankCorporaWebPage} that is representative of the distribution of real-world language usage.  
The algorithm identifies the discrepancy between the seed syntax 
and the reference grammar to decide how a seed can be expanded.
%This phase builds \pcfg from the reference corpus. \cfg
%is constructed by parsing \sents in the corpus and extracting
%\prodrs. In addition, the probability of \cfg is estimated by
%its frequency over corpus. The output of this phase is the constructed
%\pcfg, and it is compared with seed parse trees. For our
%implementation, we build the \pcfg from the Penn Treebank corpus dataset.

@@

For each production of in each seed's parse tree (lines 3 and 4), we
extract its non-terminal at the left-hand-side (line 5), $s\_lhs$, and
the grammar symbols at the right-hand-side (line 6), $s\_rhs$. In line
7, the algorithm iterates through all productions in the reference
context-free grammar and matches these that have the same non-terminal
at the left-hand-side as $s\_lhs$.  The right-hand-side of each
matched production is called $r\_rhs$.  If $s\_rhs$ consists of a
subset of the grammar symbols in $r\_rhs$ (line 8), the additional
symbols in the $r\_rhs$ are inserted as masks in the parse tree of
the seed \sent, in their respective positions in the expanded production.
The left to right traversal of the leaves of an expanded parse tree
forms a masked \sent.  Due to a potential large number of masked \sents created by this algorithm, we randomly select $k$
masked \sents for the next \sent expansion and validation phase (line 14).
% The
% random sampling is unbiased approach since it gives same chance to be
% chosen. Thus, the random sample becomes representative of the
% population of the masked \sents, and it efficiently shows the
% usefulness of the \tool.

\begin{figure*}[t]
  \centering
  \includegraphics[scale=0.7]{figs/running_example.pdf}
  \vspace{-5mm}
  \caption{\RunningExCaption}
%  Expansion of the seed \sent ``Or
%both.''. For a \prodr in seed ``NP->[DT]'' on the left, the \prodr of
%``NP->[DT, NNS]'' is found in reference. Thus, the component NNS can be expanded in the
%seed and ``NP->[DT]'' is replaced with ``NP->[DT, NNS]'' and it generates new
%expanded \sent on the right.
\end{figure*}

\paragraph{Running example.} Figure~\ref{fig:ExpEx} shows an example using Algorithm \ref{alg:diff}
to generate a masked \sent. The \sent ``Or both." is a seed of the \lc of
\SareqExOne.  The tree on the left shows the parse tree of this seed;
it consists of two productions: ``FRAG->[CC, NP, .]" and ``NP->[DT]".
When matching the left-hand-side non-terminal of the second production
(i.e., ``NP") in the reference CFG, we found that it includes a
production ``NP->[DT, NNS]" which has an additional symbol ``NNS" on
the right-hand-side.  The algorithm thus expands the parse tree with
this symbol, shown in the second tree.  The masked \sent ``Or both
\{MASK\}." is the result of the left-to-right traversal of this
expanded parse tree.

%Given the seed parse tree and reference \pcfg, production
%differentiation phase suggests structural expansion candidates on the
%seed input. this phase aims to analyze which structural components and
%where they can be added into the seed structure for its expansion. To
%do so, we explore reference \prodrs comparing it with each \prodr used
%in seed input. This results in the phase described in
%Algorithm~\ref{code:ProdDiffAlg}. For each production rule in seed
%inputs ($seed\_prod$), it searches production rules in reference
%($ref\_prod$) which it has same non-terminal on the \lhs
%($seed\_prod.lhs==ref\_prod.lhs$) and superset of \rhs of the seed
%production rule ($seed\_prod.rhs \subset ref\_prod.rhs$).  As we
%assume that the reference \cfg is built from real world data
%distribution, the elements in the complement set ($ref\_prod.rhs -
%seed\_prod.rhs$) become an expansion candidate which can be expanded
%from the $seed\_prod.rhs$ found in real world. In addition, the
%measure of how consistent the production rule is with the given seed
%structure is given in its probability of the reference production rule
%($ref\_prob$) multiplied by that of parents of $seed\_prod$. The
%expansion candidate consists of terminal or nonterminal symbols. When
%there is a phrase-level or clause-level nonterminal symbol, \eg noun
%phrase, it needs to be expanded and replaced with word-level
%nonterminal or terminal symbols to generate the expansion
%candidate. The number of feasible replacement is unbounded because of
%its high degree of freedom. Therefore, in this work, we focus on the
%expansion candidate with only the word-level nonterminal or terminal
%symbols for the effectiveness of \Model. Lastly, the expanded
%component is replaced with the mask token for the next phase. The
%example of the expansion is illustrated in
%Figure~\ref{fig:ExpEx}. The ``NP->[DT]'' is queried into
%reference, and ``NP->[DT,NNS]'' is identified as its expansion
%candidate since the \rhs of ``NP->[DT,NNS]'' is superset of that of
%``NP->[DT]''. the component of NNS is replaced with mask token in
%\sent-level. Therefore, the ``Or both \{MASK\}.'' is suggested for
%the next phase.

@@

In this phase, the words to fill in the masks in the masked \sents are
suggested by the BERT model~\cite{devlin2019bert}. 
BERT is a transformer-based \nl model that is pre-trained on two tasks: masked token prediction and next sentence prediction. BERT model is capable of suggesting words for the masked
token according to its surrounding context in a \sent. For each masked
token, multiple words are suggested, ranked by their confidence scores.
% \sw{Algorithm 1 does not require we only have one mask in the masked
%   \sent. Can the model suggest multiple words at the same time?}
% \jl{yes. it can predict multiple masked tokens at the same time.}
% \jl{I added the explanation of BERT}
%
Because BERT model is not aware of the \lc
specification and the grammar symbol in the expanded parse tree, an
expanded \sent using the suggested words may no longer satisfy the
\lc specification. Therefore, we perform validation
on the suggested words and only accept them if the following three
criteria are met.

First, the PoS tag of the suggested word must match the PoS tag of the
expanded symbol in the parse tree. For the example in
Figure~\ref{fig:ExpEx}, the masked symbol is a ``NNS" (i.e., plural
noun); thus, the suggested word must also be a ``NNS". We use \spacy \cite{spacy}, a free open-source library for \nlp, to extract the PoS
tag for each suggested word.

Second, it is required that the
sentiment of the expanded \sent is the same as the seed \sent. To
ensure this, the suggested words must be \neu.

Third, we additionally verify that the
expanded \sents satisfy the same search rules for the seed
\sent in LC1 and LC2. This criteria cannot be applied to other linguistic capabilities because they have additional transformation templates.

% In this work, out of the BERT suggested words, we randomly selects
% words that matches their PoS tag of the expanded symbol in the parse
% tree. The words that meets the second and third criteria are finally
% employed for the expanded \sents.

% Our goal for generating the expanded \sents is to use them for
% evaluating the \sa models on the associated \lc in addition to the
% seed \sent. It is only achieved when the expanded \sents are also met
% with the same search rules for the \lc. For LC1 as an example, the
% expanded \sent must still conform to the specification of the seed's
% \lc specification. Therefore, the expanded \sents are required to be
% short and to only have \neu adjectives and \nns.
% \sw{Say why we use
%   the third criteria only for LC1 and LC2.}  \jl{I added the
%   explanation}
